Tilting at Algorithmic Windmills

### 2013-05-06 Tilting/paper.tex
WIP: notes on this paper
WIP: illustrate what we have in mind by an example
ENH:  Paramore new album

\usepackage{color,hyperref}
\usepackage{multirow,booktabs,ctable,array}
\usepackage{lscape}
\usepackage{amsmath}
\usepackage{lineno}
\usepackage{ulem}
\usepackage{setspace}
\usepackage{listings}
\usepackage{float}
\usepackage{listings}
\usepackage{color,colortbl}
\usepackage{rccol}
\usepackage[table]{xcolor}

@@

\begin{abstract} 
Exploration of neuroscience hypotheses have been greatly enhanced
by the increased availability of high-performance computational resources, large-scale
communal efforts such as the Insight Toolkit and the R statistical project and
algorithmic advancements for data transformations and processing. 
An integral component of the vetting process for novel algorithmic techniques
includes comparison with other methods previously
established within the community.  Public availability, including 
open source distribution, of established packages has significantly 
facilitated these comparative evaluations.  However, complementing the
recent set of papers pointing to serious methodological and statistical
bias considerations in neuroimaging research, we point out potential 
issues associated with a type of measurement bias in comparative algorithmic
evaluations and propose a set of guidelines for authors and reviewers
to minimize this confound.
\end{abstract}
\begin{keyword}
open science \sep reproducibility 
%% keywords here, in the form: keyword \sep keyword
\end{keyword}

@@

%\input{intro}
%
%\input{methods}

\section{Introduction}

@@

  \item
    \textcolor{blue}{\href{http://en.wikipedia.org/wiki/Accuracy_and_precision}{Accuracy
        and precision (link)}} --- most studies ignore precision.  it
    is critical in longitudinal studies.   precision is impacted by
    parameter choices as well as computational architecture
    (Freesurfer paper)

@@

  \item Prediction---an algorithm that yields a ``more significant''
    result is not always better.  might mean ``finds more effected
    areas'' or smaller p-values.  Testing prediction is complements
    studies of significance and cumulative distribution functions.

>>>>

\usepackage{color}
\usepackage{multirow,booktabs,ctable,array}
\usepackage{lscape}
\usepackage{amsmath}
\usepackage{lineno}
\usepackage{ulem}
\usepackage{setspace}
\usepackage{listings}
\usepackage{float}
\usepackage{listings}
\usepackage{color,colortbl}
\usepackage{rccol}
\usepackage[table]{xcolor}

@@

\begin{abstract} 
Exploration of neuroscience hypotheses have been  enhanced
by the increased availability of high-performance computational resources, large-scale
communal efforts such as the Insight Toolkit and the R statistical project and
algorithmic advancements for data transformations and processing. 
An integral component of the vetting process for novel algorithmic techniques
includes comparison with other methods previously
established within the community.  Public availability, including 
open source distribution, of established packages has significantly 
facilitated these comparative evaluations.  However, complementing the
recent set of papers pointing to serious methodological and statistical
bias considerations in neuroimaging research, we point out potential 
issues associated with a type of measurement bias in comparative algorithmic
evaluations and propose a set of guidelines for authors and reviewers
to minimize this confound.
\end{abstract}
\begin{keyword}
open science \sep reproducibility 
%% keywords here, in the form: keyword \sep keyword
\end{keyword}

@@

\section{Introduction}
The neuroimaging community has benefited from the proliferation
of imaging software.  Established packages such as SPM from the 
Wellcome Trust Centre for Neuroimaging \cite{Ashburner2012},
the FMRIB Software Library (FSL) \cite{Jenkinson2012}, and
the AFNI toolkit \cite{cox2012} have aided neuroimaging researchers 
around the world in performing complex analyses as part of 
ongoing neuroscience research.  Parallel to this line of
inquiry is continued algorithmic innovation for improvement
in analysis tools.

As fellow scientists who actively participate in this type of 
research, we have noticed several publications in which a principle
component involves the comparison of algorithms.  One of our
concerns is the lack of detail with which these comparisons are
presented and the corresponding possibility of {\it instrumentation
bias} \cite{sackett1979} (considering software  as a type of 
instrument requiring proper ``calibration'' for accurate measurements).  
In this editorial we propose an initial set of guidelines for minimal
reporting of algorithmic usage to minimize such bias understanding that 
the discussion will hopefully 
elicit a much more comprehensive response from the larger community.

Although previous articles have discussed similar concerns, oftentimes
within a much larger context (e.g. fMRI reporting \cite{poldrack2008}), 
we are particularly interested in comparative evaluations of software
and provide additional discussion through experience and actual 
examples found within the literature.
It is hoped that this commentary serves to raise awareness to both 
authors and reviewers to such problematic issues (not unlike other
recent articles detailing additional potential sources of methodological 
bias \cite{kriegeskorte2009,vul2012,tustison2012}).

\section{Guidelines}

\subsection{Parameters should be included (perhaps the precise command line call)}
Perhaps the most blatantly egregious sin is omitting

\subsection{Hassle the developers for a set of default parameters}

\subsection{Scripts used to invoke the algorithms should be posted}

\subsection{Comparisons should be performed on publicly available data}

\subsection{Corollary to the above:  Resulting images should be publicly posted}

\subsection{Scripts used to create the plots should also be posted}

\subsection{When possible, consultation should be made with the original authors}

\subsection{Provide details of where the algorithm was obtained}

\subsection{Provide system details of where the algorithm was run}

\subsection{Avoid confirmation bias:  Don't code up your own version particularly if one
is already available}

\subsection{Co-authors should verify findings}

\subsection{Properly contextualize comparisons}
Accuracy is important but other considerations should be included such
as biological plausibility.  Torsten's CURT algorithm could produce an ``accurate''
registration but not one which is biologically plausible.

\section{Conclusion}

\section{Acknowledgments}

@@

  \item
    \textcolor{blue}%{\href{http://en.wikipedia.org/wiki/Accuracy_and_precision}{Accuracy and precision (link)}} 
        wiki link--- most studies ignore precision.  it
    is critical in longitudinal studies.   precision is impacted by
    parameter choices as well as computational architecture
    (Freesurfer paper)

@@

  \item Prediction---an algorithm that yields a ``more significant''
    result is not always better.  might mean ``finds more effected
    areas'' or smaller p-values.  Testing prediction complements
    studies of significance and cumulative distribution functions.
